### 2018-07-29 FS/overview.tex
NIL

\documentclass{article}
\usepackage{xeCJK}
\usepackage{hyperref}
\title{Overview}
\date{}
\begin{document}

\maketitle

\tableofcontents
\newpage
\part{环境配置}
\section{虚拟机}
\begin{itemize}
\item \href{https://www.howtogeek.com/117635/how-to-install-kvm-and-create-virtual-machines-on-ubuntu/}{Installing qemu-kvm}

\end{itemize}
\section{GlusterFS}

\end{document}