### 2018-08-19 FS/overview.tex
8.19

\documentclass{article}
\usepackage{xeCJK}
\usepackage{hyperref}
\usepackage{listings}
\lstset{
 columns=fixed,       
 numbers=left,                                        % 在左侧显示行号
 numberstyle=\tiny\color{gray},                       % 设定行号格式
 frame=single,                                          % 不显示背景边框
 keywordstyle=\color[RGB]{40,40,255},                 % 设定关键字颜色
 numberstyle=\footnotesize\color{darkgray},           
 commentstyle=\it\color[RGB]{0,96,96},                % 设置代码注释的格式
 stringstyle=\rmfamily\slshape\color[RGB]{128,0,0},   % 设置字符串格式
 showstringspaces=false,                              % 不显示字符串中的空格
 language=c++,                                        % 设置语言
}

\title{Overview}
\date{}
\begin{document}

\maketitle

\tableofcontents
\newpage
\part{环境配置}
\section{虚拟机}
\begin{enumerate}
\item \href{https://blogs.oracle.com/scoter/networking-in-virtualbox-v2}{VirtualBox Networking Configuration}
\begin{itemize}
\item Network Address Translation.
\item Bridged Network.
\item Internal Network.
\item Host-only Networking.
\item Port-forwarding.
\end{itemize}
\end{enumerate}
\section{VirtualBox+Centos7+GlusterFS}
\subsection{\href{https://docs.gluster.org/en/latest/Install-Guide/Common\_criteria/}{基本概念}}
\begin{itemize}
\item \textbf{trusted pool}:hosts
\item \textbf{node}:server
\item \textbf{brick}:device(filesystem)
\item \textbf{export}:mount path of the bricks on a server
\item \textbf{Gluster Volumn(Global Namespace)}:collection of bricks
\item \textbf{GNFS and kNFS}:GNFS=inline NFS server. kNFS=kernel NFS. \href{https://en.wikipedia.org/wiki/Network\_File\_System}{NFS wiki}
\item \textbf{DNS setup}:????
\end{itemize}

\subsection{\href{https://wiki.centos.org/SpecialInterestGroup/Storage/gluster-Quickstart}{Installing GlusterFS on CentOs}}
\begin{enumerate}
\item Step 1: nodes.
\item Step 2: \href{https://gluster-documentations.readthedocs.io/en/latest/Administrator\%20Guide/Setting\%20Up\%20Volumes/}{Format and mount the bricks.}
\item Step 3: Installing glusterFS.
\item Step 4: \href{https://wiki.centos.org/HowTos/Network/IPTables}{Configure iptables.}
\item Step 5: \href{https://docs.gluster.org/en/v3/Administrator\%20Guide/Storage\%20Pools/}{Manage Trusted Storage Pool}
\item Step 6: \href{https://gluster-documentations.readthedocs.io/en/latest/Administrator\%20Guide/Setting\%20Up\%20Volumes/}{Set up a GlusterFS Volume.}
\end{enumerate}
\part{GlusterFS开发}
\href{https://docs.gluster.org/en/latest/Developer-guide/Developers-Index/\#contributing-to-the-gluster-community}{官方开发文档}
\section{Workflow}
\subsection*{\href{https://github.com/gluster/glusterfs}{Git}}
\subsection*{\href{review.gluster.org}{Gerrit}}
Code Review System.

%\subsubsection{\href{https://docs.gluster.org/en/latest/Install-Guide/compiling-rpms/}{如何编译GlusterFS}}

\part{文献}
\section{GlusterFS}
\end{document}